---
title: "Introduction to RSQLite"
author: "Yang Yang"
date: "November 3, 2014"
output: 
  ioslides_presentation:
    css: custom.css
---
```{r setup, echo=F,include=F}
knitr::opts_chunk$set(comment = NA, highlight=T)
```

## What is SQL


- Structured Query Language
- Relational Database Management System (RDBMS)


## SQLite vs SQL

- "Lite" verison of SQL
- Supports most of the SQL syntax
- Only allow single writer at a time
- No user management
- Best for mobile applications, testing
- **NOT** for big-scale data
- **NOT** for enterprises due to security reasons


## Basic Definitions

- Table: collection of related data entries
- Field: column names, eqv to header in `R`
- Column
- Row
- Supported data types: **NULL**, **INTEGER**, **TEXT**, **BLOB**

## RSQLite: Installation
```{r,eval=FALSE}
install.packages(RSQLite, dependencies = T)
```

- The source includes SQLite engine
- Depends on DBI package


## RSQLite: Create a table
```{r,warning=FALSE,message=FALSE,comment = NA,results='hide'}
library(RSQLite)
db = dbConnect(SQLite(), dbname="Test.sqlite")
dbSendQuery(conn = db,
       "CREATE TABLE BASEBALL
       (Team_ID INTEGER,
         Team_Name TEXT,
        Leage TEXT,
        Payroll REAL,
        Wins INTEGER)")
```


## RSQLite: Enter data manually:
```{r,cache=F,results='hide',comment = NA}
dbSendQuery(conn = db,
         "INSERT INTO BASEBALL
         VALUES (1, 'Twins', 'American League', '54641175','1020')")
dbSendQuery(conn = db,
         "INSERT INTO BASEBALL
         VALUES (2, 'Giants', 'American League', '82288960','1033')")
dbSendQuery(conn = db,
         "INSERT INTO BASEBALL
         VALUES (3, 'Royals', 'National League', '49816557','803')")
```


## RSQLite: Check the content of the table
```{r,comment = NA}
## The tables in the database
dbListTables(db) 
## The columns in a table
dbListFields(db, "BASEBALL")   
## The data in a table
head(dbReadTable(db, "BASEBALL"))
```

```{r,echo=FALSE,include=FALSE}
dbRemoveTable(db, "BASEBALL")
```



## RSQLite: Read data from csv file

- Windows and UNIX uses different End of Line
- Always use `read.csv` to read csv into `R` dataframe first
```{r,comment=F,message=F,warning=F,tidy=F,results='hide',eval=F}
bball = read.csv('bball.csv')
db.bball = dbConnect(SQLite(), dbname="bball.sqlite")
dbWriteTable(conn = db.bball, name = "BASEBALL", bball, overwrite=T,
             row.names=FALSE)
```
```{r,echo=FALSE,include=F}
db.bball = dbConnect(SQLite(), dbname="bball.sqlite")
```

```{r,comment=F,message=F,warning=F,tidy=TRUE}
tmp=dbReadTable(db.bball, "BASEBALL")
```
```{r,echo=F}
rownames(tmp)=NULL
```
```{r,echo=F}
tmp
```


## RSQLite: make queries
- A query: which teams are in National League?
- Select all columns
```{r}
dbGetQuery(db.bball, "select * from BASEBALL where League='NL'")[1:2,]
```
- Select some columns
```{r,tidy=FALSE}
dbGetQuery(db.bball, "select League, Team, RSW from BASEBALL
           where League='NL'")[1:2,]
```

## RSQLite: more complicated queries
- SQL logical operators: AND, OR, NOT
- which American League teams has regular season wins less than 810?
```{r,tidy=FALSE}
dbGetQuery(db.bball, "select * from BASEBALL where League='AL' and RSW<810")

```


